### 2017-12-19 paper/dispRity_paper_highlighted.tex
Updated paper

\begin{table}
\resizebox{\textwidth}{!}
{%
    \begin{tabular}{p{2cm}|p{5cm}|p{1cm}|p{5cm}|p{5cm}}
        name & description & dim level & definition & source \\
        \hline
        \texttt{ancestral. dist} & The distance between an element and its ancestor & 2 & $\sqrt{\sum_{i=1}^{n}{({k}_{n}-Ancestor_{n})^2}}$ & This package \\
        \texttt{centroids} & The distance between each element and a fixed point$^1$ of the space & 2 & $\sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}$ & This package \\
        \texttt{convhull. surface} & The surface of the convex hull & 1 & NA & \texttt{geometry::convhulln} \\
        \texttt{convhull. volume} & The volume of the convex hull & 1 & NA & \texttt{geometry::convhulln} \\
        \texttt{diagonal} & The greatest Euclidean distance & 1 & $\sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}$ & This package \\
        \texttt{ellipse. volume}$^2$ & The volume of the ellipsoid & 1 & $\frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})$ & \cite{DonohueDim}\\
        \texttt{mode.val} & The modal value & 1 & NA & This package\\
        \texttt{n.ball. volume} & The hyper-spherical ($n$-ball) volume & 1 & $\frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} R$ & This package\\
        \texttt{pairwise. dist} & The pairwise distances between elements & 2 & NA & \texttt{stats::dist}\\        
        \texttt{radius} & The radius of each dimensions & 2 & $|\frac{\sum_{i=1}^{n}k_i}{n} - f(\mathbf{v}k)|$ & This package\\
        \texttt{ranges} & The absolute ranges of each dimension & 2 & $|max(k_i) - min(k_i)|$ & This package \\
        \texttt{span.tree. length} & The minimal spanning tree length & 1 & $\sum(\mathrm{branch\ length})$ & \texttt{vegan::spantree}\\        
        \texttt{variances} & The variance of each dimension & 2 & $\sigma^{2}{k_i}$ & This package \\
    \end{tabular}
}%
    % \caption{\small{Where $k$ is the number of dimensions, $n$ the number of elements, $\Gamma$ is the Gamma distribution, $\lambda_i$ is the eigenvalue of each dimensions, $\sigma^{2}$ is their variance and $Centroid_{k}$ is their mean, $Ancestor_{n}$ is the coordinates of the ancestor of element $n$, $f(\mathbf{v}k)$ is function to select one value from the vector $\mathbf{v}$ of the dimension $k$ (e.g. it's maximum, minimum, mean, etc.), $R$ is the radius of the sphere or the product of the radii of each dimensions ($\displaystyle\prod_{i=1}^{k}$ - for a hyper-ellipsoid). $^1$ by default that point is the centroid of the elements; $^2$ this function uses a fast estimation of the eigenvalue that only works in an ordinated space based on MDS or PCO/PCoA (\textit{not} PCA).}}
    % \label{Tab:metrics}
\end{table}

@@

Some tests are implemented within the package such as the Bhattacharrya Coefficient \citep[\texttt{bhatt.coeff};][]{Bhattacharyya,GuillermeCooper} or a permutation test based on null hypothesised multidimensional space following \cite{diaz2016global} (\texttt{null.test}).
Additionally, this function also allows additional arguments such as \texttt{rarefaction} (as described above) or \texttt{correction} to adjust p-values when using multiple parametric tests.

@@

From the plot and the summary, disparity seems to change (increase) through time from the Cretaceous to the Eocene.
Similarly to the example above, it is also possible to statistically test this hypothesis using, for example, an ANOVA (Table \ref{Tab:anova}):

\noindent \texttt{> anova(test.dispRity(disparity2, test = lm, comparisons = \textquotedbl all\textquotedbl))}

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrr}
  \hline
 & Df & Sum Sq & Mean Sq & F value & Pr($>$F) \\ 
  \hline
subsets & 2 & 3.14 & 1.57 & 177.84 & 0.0000 \\ 
  Residuals & 297 & 2.62 & 0.01 &  &  \\ 
   \hline
\end{tabular}
\caption{ANOVA performed on a \disp object.}
\label{Tab:anova}
\end{table}

@@

\section{Acknowledgments}
Many thanks to Natalie Cooper for encouraging the development of this package and helping with the writing of this paper and the package manuals.
Thanks to Martin Brazeau and Andrew Jackson for specific implementation ideas 
\hl{and to Michael Collyer, Gavin Simpson and two other anonymous reviewers for constructive comments on the earlier version of this manuscript}. %TG: add Graeme Lloyd, Dave Bapst and Emma Sherratt
I acknowledge support from European Research Council under the European Union's Seventh Framework Programme (FP/2007 – 2013)/ERC Grant Agreement number 311092 awarded to Martin D. Brazeau 
\hl{and from the Discovery Project Grant DP170103227 awarded to Vera Weisbecker}.

>>>>

\begin{table}
\resizebox{\textwidth}{!}
{%
    \begin{tabular}{p{2cm}|p{5cm}|p{1cm}|p{5cm}|p{5cm}}
        name & description & dim & definition & source \\
        \hline
        \texttt{ancestral. dist} & The distance between an element and its ancestor & 2 & $\sqrt{\sum_{i=1}^{n}{({k}_{n}-Ancestor_{n})^2}}$ & This package \\
        \texttt{centroids} & The distance between each element and a fixed point$^1$ of the space & 2 & $\sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}$ & This package \\
        \texttt{convhull. surface} & The surface of the convex hull & 1 & NA & \texttt{geometry::convhulln} \citep{barber1996quickhull,geometryR}\\
        \texttt{convhull. volume} & The volume of the convex hull & 1 & NA & \texttt{geometry::convhulln} \citep{barber1996quickhull,geometryR}\\
        \texttt{diagonal} & The greatest Euclidean distance & 1 & $\sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}$ & This package \\
        \texttt{ellipse. volume}$^2$ & The volume of the ellipsoid & 1 & $\frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})$ & This package; based on \cite{DonohueDim}\\
        \texttt{mode.val} & The modal value & 1 & NA & This package\\
        \texttt{n.ball. volume} & The hyper-spherical ($n$-ball) volume & 1 & $\frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} R$ & This package\\
        \texttt{pairwise. dist} & The pairwise distances between elements & 2 & NA & \texttt{vegan::vegdist} \citep{oksanen2007vegan}\\        
        \texttt{radius} & The radius of each dimensions & 2 & $|\frac{\sum_{i=1}^{n}k_i}{n} - f(\mathbf{v}k)|$ & This package\\
        \texttt{ranges} & The absolute ranges of each dimension & 2 & $|max(k_i) - min(k_i)|$ & This package \\
        \texttt{span.tree. length} & The minimal spanning tree length & 1 & $\sum(\mathrm{branch\ length})$ & \texttt{vegan::spantree} \citep{oksanen2007vegan}\\
        \texttt{variances} & The variance of each dimension & 2 & $\sigma^{2}{k_i}$ & This package \\
    \end{tabular}
}%
    % \caption{\small{Where $k$ is the number of dimensions, $n$ the number of elements, $\Gamma$ is the Gamma distribution, $\lambda_i$ is the eigenvalue of each dimensions, $\sigma^{2}$ is their variance and $Centroid_{k}$ is their mean, $Ancestor_{n}$ is the coordinates of the ancestor of element $n$, $f(\mathbf{v}k)$ is function to select one value from the vector $\mathbf{v}$ of the dimension $k$ (e.g. it's maximum, minimum, mean, etc.), $R$ is the radius of the sphere or the product of the radii of each dimensions ($\displaystyle\prod_{i=1}^{k}$ - for a hyper-ellipsoid). $^1$ by default that point is the centroid of the elements; $^2$ this function uses a fast estimation of the eigenvalue that only works in an ordinated space based on MDS or PCO/PCoA (\textit{not} PCA).}}
    % \label{Tab:metrics}
\end{table}

@@

Some tests are implemented within the package such as the Bhattacharrya Coefficient \citep[\texttt{bhatt.coeff};][]{Bhattacharyya,GuillermeCooper}, a permutation test based on null hypothesised multidimensional space following \cite{diaz2016global} (\texttt{null.test}) 
\hl{as well as a wrapper for the \texttt{vegan::adonis}}
 \citep{oksanen2007vegan}
\hl{and \texttt{geiger::dtt}}
 \citep{geiger2008}
\hl{functions (respectively \texttt{adonis.dispRity} and \texttt{dtt.dispRity} - allowing \texttt{dispRity} objects to be directly used with any disparity metric).}
Additionally, this function also allows additional arguments such as \texttt{rarefaction} (as described above) or \texttt{correction} to adjust p-values when using multiple parametric tests.

@@

From the plot and the summary, disparity seems to change (increase) through time from the Cretaceous to the Eocene.
Similarly to the example above, it is also possible to statistically test this hypothesis using, for example, permutational multivariate ANOVA \citep[PERMANOVA; ][]{permanova} through the \texttt{adonis.disprity} function that is a wrapper of the \texttt{vegan::adonis} function \citep{oksanen2007vegan} for \disp objects:

\noindent \texttt{> test.dispRity(disparity2, test = adonis.dispRity)}

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
              & Df & Sum Sq & Mean Sq & F Model & R2   & Pr($>$F) \\ 
  \hline
    time      & 2  &  7.50 & 3.75     & 2.06    & 0.11 & <0.01 *** \\ 
    Residuals & 34 & 61.82 & 1.81     &         & 0.89 & \\ 
    Total     & 36 & 69.32 &          &         & 1.00 & \\ 
   \hline
\end{tabular}
\caption{Permanova output from \texttt{adonis.dispRity}:\\
Call: vegan::adonis(formula = dist(matrix) \texttildelow time, data = disparity2, method = "euclidean"); Permutation: free; Number of permutations: 999; Terms added sequentially (first to last). Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1 ` ' 1.}
\label{Tab:anova}
\end{table}

@@

\section{Acknowledgments}
Many thanks to Natalie Cooper for encouraging the development of this package and helping with the writing of this paper and the package manuals.
Thanks to David Bapst, Martin Brazeau, Andrew Jackson, Graeme Lloyd and Emma Sherratt for specific implementation ideas 
\hl{and to Michael Collyer, Gavin Simpson and two other anonymous reviewers for constructive and encouraging comments on the manuscript}. %TG: add Graeme Lloyd, Dave Bapst and Emma Sherratt
I acknowledge support from European Research Council under the European Union's Seventh Framework Programme (FP/2007 – 2013)/ERC Grant Agreement number 311092 awarded to Martin D. Brazeau 
\hl{and from the Discovery Project Grant DP170103227 awarded to Vera Weisbecker}.
